---
title: "Preprocessing for Stan Analysis"
author: "Asger"
date: "11/9/2024"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(pacman, tidyverse, stringr, posterior, cmdstanr)

source("src/Coherence Analysis Functions.R")
```



```{r Data Cleanup}

CoherenceLong = read_csv("temp/CoherenceLong.csv")



CoherenceLong$pair =CoherenceLong$pair %>% as.factor() %>% factor(levels=c('1', '2', '3'
                                                                ,
                                                                '4',
                                                                "5",
                                                                "6",
                                                                "7",
                                                                "8",
                                                                "9",
                                                                "10",
                                                                "11",
                                                                "12",
                                                                "13",
                                                                "14",
                                                                "16",
                                                                "17",
                                                                "18",
                                                                "19",
                                                                "20",
                                                                "21",
                                                                "22",
                                                                "23",
                                                                "Surrogate1",
                                                                "Surrogate2",
                                                                "Surrogate3",
                                                                "Surrogate4",
                                                                "Surrogate5",
                                                                "Surrogate6",
                                                                "Surrogate7",
                                                                "Surrogate8",
                                                                "Surrogate9",
                                                                "Surrogate10",
                                                                "Surrogate11",
                                                                "Surrogate12",
                                                                "Surrogate13",
                                                                "Surrogate14",
                                                                "Surrogate15",
                                                                "Surrogate16",
                                                                "Surrogate17",
                                                                "Surrogate18",
                                                                "Surrogate19",
                                                                "Surrogate20",
                                                                "Surrogate21",
                                                                "Surrogate22"))




CoherenceLong$visit = CoherenceLong$visit %>% as.factor() %>% factor(levels = c(
  "1", "4"
)) # 1,4 -> 1, 2 -> 0, 1



CoherenceLong$pair = dummy_code(CoherenceLong$pair)  # First 22 are real, last 22 are surrogate
CoherenceLong$visit = dummy_code(CoherenceLong$visit) # 1 is 1, 2 is 4

CoherenceLong$surrogate =  ifelse(as.numeric(CoherenceLong$pair) < 23, 0, 1)

CoherenceLong$condition = ifelse(CoherenceLong$condition == "rest_avg", 0, 1)

all_channels = c(
  'S1_D1 hbo', 'S1_D1 hbr', 'S1_D2 hbo', 'S1_D2 hbr', 'S1_D3 hbo', 'S1_D3 hbr', 'S2_D1 hbo', 'S2_D1 hbr', 'S2_D4 hbo', 'S2_D4 hbr', 'S3_D1 hbo', 'S3_D1 hbr', 'S3_D3 hbo', 'S3_D3 hbr', 'S3_D4 hbo', 'S3_D4 hbr', 'S4_D4 hbo', 'S4_D4 hbr', 'S4_D5 hbo', 'S4_D5 hbr', 'S4_D10 hbo', 'S4_D10 hbr', 'S5_D2 hbo', 'S5_D2 hbr', 'S5_D3 hbo', 'S5_D3 hbr', 'S5_D7 hbo', 'S5_D7 hbr', 'S6_D2 hbo', 'S6_D2 hbr', 'S6_D7 hbo', 'S6_D7 hbr', 'S7_D3 hbo', 'S7_D3 hbr', 'S7_D5 hbo', 'S7_D5 hbr', 'S7_D6 hbo', 'S7_D6 hbr', 'S7_D8 hbo', 'S7_D8 hbr', 'S8_D6 hbo', 'S8_D6 hbr', 'S8_D7 hbo', 'S8_D7 hbr', 'S8_D16 hbo', 'S8_D16 hbr', 'S9_D4 hbo', 'S9_D4 hbr', 'S9_D9 hbo', 'S9_D9 hbr', 'S10_D4 hbo', 'S10_D4 hbr', 'S10_D9 hbo', 'S10_D9 hbr', 'S10_D10 hbo', 'S10_D10 hbr', 'S10_D11 hbo', 'S10_D11 hbr', 'S11_D5 hbo', 'S11_D5 hbr', 'S11_D8 hbo', 'S11_D8 hbr', 'S11_D10 hbo', 'S11_D10 hbr', 'S12_D9 hbo', 'S12_D9 hbr', 'S12_D11 hbo', 'S12_D11 hbr', 'S12_D12 hbo', 'S12_D12 hbr', 'S13_D10 hbo', 'S13_D10 hbr', 'S13_D11 hbo', 'S13_D11 hbr', 'S13_D13 hbo', 'S13_D13 hbr', 'S14_D11 hbo', 'S14_D11 hbr', 'S14_D12 hbo', 'S14_D12 hbr', 'S14_D13 hbo', 'S14_D13 hbr', 'S14_D14 hbo', 'S14_D14 hbr', 'S15_D12 hbo', 'S15_D12 hbr', 'S15_D14 hbo', 'S15_D14 hbr', 'S15_D15 hbo', 'S15_D15 hbr', 'S16_D13 hbo', 'S16_D13 hbr', 'S16_D14 hbo', 'S16_D14 hbr', 'S17_D7 hbo', 'S17_D7 hbr', 'S17_D17 hbo', 'S17_D17 hbr', 'S18_D6 hbo', 'S18_D6 hbr', 'S18_D8 hbo', 'S18_D8 hbr', 'S18_D16 hbo', 'S18_D16 hbr', 'S19_D7 hbo', 'S19_D7 hbr', 'S19_D16 hbo', 'S19_D16 hbr', 'S19_D17 hbo', 'S19_D17 hbr', 'S19_D18 hbo', 'S19_D18 hbr', 'S20_D16 hbo', 'S20_D16 hbr', 'S20_D18 hbo', 'S20_D18 hbr', 'S20_D19 hbo', 'S20_D19 hbr', 'S21_D17 hbo', 'S21_D17 hbr', 'S21_D18 hbo', 'S21_D18 hbr', 'S21_D20 hbo', 'S21_D20 hbr', 'S22_D18 hbo', 'S22_D18 hbr', 'S22_D19 hbo', 'S22_D19 hbr', 'S22_D20 hbo', 'S22_D20 hbr', 'S22_D21 hbo', 'S22_D21 hbr', 'S23_D19 hbo', 'S23_D19 hbr', 'S23_D21 hbo', 'S23_D21 hbr', 'S24_D20 hbo', 'S24_D20 hbr', 'S24_D21 hbo', 'S24_D21 hbr', 'S24_D22 hbo', 'S24_D22 hbr'
)

hbo_channels = c(
  'S1_D1 hbo',  'S1_D2 hbo',  'S1_D3 hbo',  'S2_D1 hbo', 'S2_D4 hbo',  'S3_D1 hbo',  'S3_D3 hbo',  'S3_D4 hbo',  'S4_D4 hbo', 'S4_D5 hbo',  'S4_D10 hbo', 'S5_D2 hbo',  'S5_D3 hbo', 'S5_D7 hbo',  'S6_D2 hbo',  'S6_D7 hbo',  'S7_D3 hbo',  'S7_D5 hbo',  'S7_D6 hbo',  'S7_D8 hbo', 'S8_D6 hbo',  'S8_D7 hbo',  'S8_D16 hbo',  'S9_D4 hbo', 'S9_D9 hbo',  'S10_D4 hbo',  'S10_D9 hbo',  'S10_D10 hbo', 'S10_D11 hbo',  'S11_D5 hbo',  'S11_D8 hbo',  'S11_D10 hbo',  'S12_D9 hbo',  'S12_D11 hbo', 'S12_D12 hbo',  'S13_D10 hbo',  'S13_D11 hbo',  'S13_D13 hbo', 'S14_D11 hbo', 'S14_D12 hbo', 'S14_D13 hbo',  'S14_D14 hbo',  'S15_D12 hbo',  'S15_D14 hbo',  'S15_D15 hbo',  'S16_D13 hbo', 'S16_D14 hbo',  'S17_D7 hbo',  'S17_D17 hbo',  'S18_D6 hbo', 'S18_D8 hbo',  'S18_D16 hbo',  'S19_D7 hbo',  'S19_D16 hbo',  'S19_D17 hbo', 'S19_D18 hbo', 'S20_D16 hbo',  'S20_D18 hbo',  'S20_D19 hbo',  'S21_D17 hbo',  'S21_D18 hbo', 'S21_D20 hbo',  'S22_D18 hbo',  'S22_D19 hbo', 'S22_D20 hbo',  'S22_D21 hbo',  'S23_D19 hbo',  'S23_D21 hbo',  'S24_D20 hbo',  'S24_D21 hbo', 'S24_D22 hbo')


hbr_channels = c('S1_D1 hbr',  'S1_D2 hbr',  'S1_D3 hbr',  'S2_D1 hbr', 'S2_D4 hbr',  'S3_D1 hbr',  'S3_D3 hbr',  'S3_D4 hbr',  'S4_D4 hbr', 'S4_D5 hbr',  'S4_D10 hbr', 'S5_D2 hbr',  'S5_D3 hbr', 'S5_D7 hbr',  'S6_D2 hbr',  'S6_D7 hbr',  'S7_D3 hbr',  'S7_D5 hbr',  'S7_D6 hbr',  'S7_D8 hbr', 'S8_D6 hbr',  'S8_D7 hbr',  'S8_D16 hbr',  'S9_D4 hbr', 'S9_D9 hbr',  'S10_D4 hbr',  'S10_D9 hbr',  'S10_D10 hbr', 'S10_D11 hbr',  'S11_D5 hbr',  'S11_D8 hbr',  'S11_D10 hbr',  'S12_D9 hbr',  'S12_D11 hbr', 'S12_D12 hbr',  'S13_D10 hbr',  'S13_D11 hbr',  'S13_D13 hbr', 'S14_D11 hbr', 'S14_D12 hbr', 'S14_D13 hbr',  'S14_D14 hbr',  'S15_D12 hbr',  'S15_D14 hbr',  'S15_D15 hbr',  'S16_D13 hbr', 'S16_D14 hbr',  'S17_D7 hbr',  'S17_D17 hbr',  'S18_D6 hbr', 'S18_D8 hbr',  'S18_D16 hbr',  'S19_D7 hbr',  'S19_D16 hbr',  'S19_D17 hbr', 'S19_D18 hbr', 'S20_D16 hbr',  'S20_D18 hbr',  'S20_D19 hbr',  'S21_D17 hbr',  'S21_D18 hbr', 'S21_D20 hbr',  'S22_D18 hbr',  'S22_D19 hbr', 'S22_D20 hbr',  'S22_D21 hbr',  'S23_D19 hbr',  'S23_D21 hbr',  'S24_D20 hbr',  'S24_D21 hbr', 'S24_D22 hbr')


CoherenceHbo = CoherenceLong %>% filter(channel %in% hbo_channels)

CoherenceHbr = CoherenceLong %>% filter(channel  %in% hbr_channels)


CoherenceHbo$channel = CoherenceHbo$channel  %>% as.factor() %>% factor(levels = hbo_channels)

CoherenceHbr$channel = CoherenceHbr$channel  %>% as.factor() %>% factor(levels = hbr_channels)

DummyCodedChannels = data.frame("Montage Designation" = levels(CoherenceHbo$channel), "Dummy Code" = levels(dummy_code(CoherenceHbo$channel)))

write.csv(DummyCodedChannels, "temp/DummyCodedChannels.csv")


CoherenceHbo$channel = dummy_code(CoherenceHbo$channel) 
CoherenceHbr$channel = dummy_code(CoherenceHbr$channel)


CoherenceHbr %>% select(surrogate, pair, channel, visit, condition, coherence) %>% write_csv("temp/CoherenceHbr.csv")

CoherenceHbo %>% select(surrogate, pair, channel, visit, condition, coherence) %>% write_csv("temp/CoherenceHbo.csv")


```


Due to arima errors in the coherence estimation, we have:

HBR
Missing Visit 2 Hbr Data from:

Real Pairs:
- Pair 3, channel 66 and 68.
- pair 8, channel 29.

Surrogate pairs:
- Pair 25, channels 66 and 68.
- Pair 30, channel 29.








HBO
Missing Visit 2 Hbo Data from:

Real Pairs:
- Pair 3, channels 66 and 68

Surrogate pairs:
- Pair 25, channels 66, 68, and 70


Coherence for these channels is set to EXACTLY ZERO.
An if statement in the stan model means it will not
try to model these channels.


```{r Data Cleanup Continued}
# Identify missing data
CoherenceHbo = read_csv("temp/CoherenceHbo.csv")
CoherenceHbr = read_csv("temp/CoherenceHbr.csv")

CoherenceHbr %>% group_by(pair, channel) %>% summarize(visit, surrogate, n() )

CoherenceHbo %>% group_by(pair, channel) %>% summarize(visit, surrogate, n() )


# Manually create missing data, setting coherence to 0.

missing_data_hbr = data.frame(
  "surrogate" = c(0,0,0,0,0,0, 1,1,1,1,1,1),
  "pair" = c(3,3,3,3,8,8,25,25,25,25, 30, 30),
  channel = c(66,66,68,68,29,29, 66,66,68,68,29,29),
  visit = c(rep(2, 12)),
  condition = c(rep(c(0,1),6)),
  coherence = c(rep(0,12)))



missing_data_hbo = data.frame(
  "surrogate" = c(0,0,0,0,1,1,1,1,1,1),
  "pair" = c(3,3,3,3,25,25,25,25,25,25),
  channel = c(66,66,68,68,66,66,68,68,70,70),
  visit = c(rep(2, 10)),
  condition = c(rep(c(0,1),5)),
  coherence = c(rep(0,10))
  
  
)

# Rbinding missing data

padded_data_hbo = rbind(CoherenceHbo, missing_data_hbo)

padded_data_hbr = rbind(CoherenceHbr, missing_data_hbr)

padded_data_hbo = padded_data_hbo %>% arrange(pair, channel, condition, visit)


padded_data_hbr = padded_data_hbr %>% arrange(pair, channel, condition, visit)



# Sanity check

sanity_check = function(dataset){
  
N_channels = as.numeric(length(unique(dataset$channel)))
N_visits = as.numeric(length(unique(dataset$visit)))
N_pairs = as.numeric(length(unique(dataset$pair)))
N_pairs_real = N_pairs/2
N_pairs_surrogate = N_pairs/2
N_conditions = as.numeric(length(unique(dataset$condition)))


data_array = array(data = NA, dim = c(N_visits, N_conditions, N_channels, N_pairs))

row = 0
  for (p in 1:N_pairs){
  for (ch in 1:N_channels){
  for (c in 1:N_conditions){ 
  for (v in 1:N_visits){
    row = row +1
    data_array[v,c,ch, p] = dataset$coherence[row]
    
  }}}}
  print("Are there any missing rows?")
  print(any(is.na(data_array)))
  print("Sum of missing rows:")
  print(sum(is.na(data_array)))
}


sanity_check(padded_data_hbo)
sanity_check(padded_data_hbr)

write_csv(padded_data_hbo, "temp/CoherenceHboPadded.csv")
write_csv(padded_data_hbr, "temp/CoherenceHbrPadded.csv")



```

```{r Filter out irrelevant channels}

padded_data_hbo = read_csv("temp/CoherenceHboPadded.csv")
padded_data_hbr = read_csv("temp/CoherenceHbrPadded.csv")

# based on literature, we've selected the following channels to actually be relevant:

relevant_channels = c("S1_D1",
"S1_D2",
"S1_D3",
"S2_D1",
"S2_D4",
"S3_D1",
"S3_D3",
"S3_D4",
"S4_D4",
"S4_D5",
"S4_D10",
"S5_D2",
"S5_D3",
"S5_D7",
"S6_D2",
"S6_D7",
"S7_D3",
"S7_D5",
"S7_D6",
"S8_D6",
"S8_D7",
"S8_D16",
"S9_D4",
"S10_D4",
"S11_D5",
"S14_D13",
"S14_D14",
"S16_D13",
"S18_D6",
"S19_D7",
"S22_D19",
"S22_D21",
"S23_D19")




relevant_channels_hbr =  paste(relevant_channels, " hbr", sep = "")
relevant_channels_hbo = paste(relevant_channels, " hbo", sep = "")



DummyCodedChannels = read_csv("temp/DummyCodedChannels.csv")

relevant_channels_dummy = (DummyCodedChannels %>% filter(Montage.Designation %in% relevant_channels_hbo))$Dummy.Code


relevant_hbo = padded_data_hbo %>% filter(channel %in% relevant_channels_dummy)

relevant_hbr = padded_data_hbr %>% filter(channel  %in% relevant_channels_dummy)

write_csv(relevant_hbo, "temp/RelevantHbo.csv")
write_csv(relevant_hbr, "temp/RelevantHbr.csv")


```







```{r Data Exploration to find sensible priors}

RelevantHbo = read_csv("temp/RelevantHbo.csv") %>% filter(coherence != 0)
RelevantHbr = read_csv( "temp/RelevantHbr.csv") %>% filter(coherence != 0)
p_load(LaplacesDemon)

explore_data = function(dataset){
  
print("Mean coherence:")
mean(c(dataset$coherence)) %>% print() 
print("Standard deviation of coherence:")
sd(dataset$coherence)%>% print()  
print("Range of coherence:")
range(dataset$coherence)%>% print() 


print("Mean logit-transformed coherence:")
mean(logit(c(dataset$coherence))) %>% print() #-0.8357877
print("Standard deviation of logit-transformed coherence:")
sd(logit(dataset$coherence)) %>% print() #0.3587508
print("Range of logit-transformed coherence:")
range(logit(dataset$coherence)) %>% print() # -2.6318984  0.8363457

}


ggplot(RelevantHbo, aes(x=logit(coherence))) + geom_histogram() 


ggplot(RelevantHbo, aes(x=logit(coherence))) + geom_histogram() + facet_wrap(~pair)


```

```{r}

explore_data(RelevantHbo)

# 
# ggplot(RelevantHbo, aes(x=logit(coherence))) + geom_histogram() 
# 
# 
# ggplot(RelevantHbo, aes(x=logit(coherence))) + geom_histogram() + facet_wrap(~surrogate)
# ggplot(RelevantHbo, aes(x=logit(coherence))) + geom_histogram() + facet_wrap(~visit)
# ggplot(RelevantHbo, aes(x=logit(coherence))) + geom_histogram() + facet_wrap(~condition)


ggplot(RelevantHbo)  +
  geom_histogram(aes(x=logit(coherence)), binwidth = 0.05, color = "black", fill = "blue", alpha = 0.3) +
  geom_vline(xintercept = 0, size = 2) +
  xlab("Coherence") +
  ylab("Count")+ xlim(c(-3, 1)) +
  theme_classic()+ ggtitle("Oxyhaemoglobin Coherence Data Distribution (Logit Transformed)")
paste("Plots", "/", "OxyHistogram", ".png", sep = "") %>% ggsave()


```
Mean coherence:
  0.3075459

Standard deviation of coherence:
  0.07374057

Range of coherence:
  0.06711349 0.69769503

Mean logit-transformed coherence:
  -0.8357877

Standard deviation of logit-transformed coherence:
  0.3587508

Range of logit-transformed coherence:
  -2.6318984  0.8363457

```{r}
explore_data(RelevantHbr)
# 
# 
# 
# ggplot(RelevantHbr, aes(x=logit(coherence))) + geom_histogram() 
# 
# 
# ggplot(RelevantHbr, aes(x=logit(coherence))) + geom_histogram() + facet_wrap(~surrogate)
# ggplot(RelevantHbr, aes(x=logit(coherence))) + geom_histogram() + facet_wrap(~visit)
# ggplot(RelevantHbr, aes(x=logit(coherence))) + geom_histogram() + facet_wrap(~condition)


ggplot(RelevantHbr)  +
  geom_histogram(aes(x=logit(coherence)), binwidth = 0.05, color = "black", fill = "blue", alpha = 0.3) +
  geom_vline(xintercept = 0, size = 2) +
  xlab("Coherence") +
  ylab("Count") + xlim(c(-3, 1)) +
  theme_classic()+ ggtitle("Deoxyhaemoglobin Coherence Data Distribution (Logit Transformed)")
paste("Plots", "/", "DeoxyHistogram", ".png", sep = "") %>% ggsave()


```
Mean coherence:
  0.303053

Standard deviation of coherence:
  0.07643093

Range of coherence:
  0.08681726 0.69760416

Mean logit-transformed coherence:
  -0.8599218

Standard deviation of logit-transformed coherence:
  0.3757122

Range of logit-transformed coherence:
  -2.3531305  0.8359149
  
  
```{r}




set.seed(1337)
ggplot(data.frame("Gamma" = c(sample(rgamma(1000,1,3),1000))))  +
  geom_histogram(aes(x=Gamma), binwidth = 0.05, color = "black", fill = "red", alpha = 0.3) +
  xlab("Sample Value") +
  ylab("Count")+ 
  theme_classic()+ ggtitle("1000 Random Samples from Gamma Distribution")
paste("Plots", "/", "Gamma_Distribution_Plot", ".png", sep = "") %>% ggsave()





```

